---
title: "the_mind"
author: "Sebastian Deri"
date: "11/5/2020"
output: html_document
---
# Settings

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. MAKE FUNCTION

```{r}
mind_prob <- function(your_hand,
                      last_pile,
                      other_n,
                      sims_n,
                      print_iter) {
  
  # 1. INFO
  # 1.1 your lowest card
  your_lowest <- min(your_hand)
  # 1.2. possible cards that could be in other player(s)' hand(s)
  
  # cards to get rid of, based on last card on pile
  if (last_pile == 0){
    get_rid <- 0
  } else if (last_pile >= 1) {
    get_rid <- last_pile:1
  }
  
  possible_cards <- setdiff(1:100, c(your_hand, get_rid))
  
  # info to display, if testing
  if(print_iter == TRUE){
    print("possible_cards:")
    print(possible_cards)
    print("~~~~~~~~~~~~~~~~~~~~~~~~~~")
  }

  # 2. CALCULATE PROBABILITIES
  # 2.1. first, get rid of impossible case #1: 
  # -- there are more cards in total (your hand + other player(s) hand(s)) than remain (# from last pile card to 100)
  if ( (length(your_hand) + other_n) > length((last_pile+1):100)){
    return(NA)
  }
  # 2.X. next, get rid of lost round #2:
  # -- if any card in your hand is lower than last card on pile
  else if (any(your_hand <= last_pile)){
    return(NA)
  }
  # 2.3. prob of lowest card, if no other player(s) have card(s)
  else if (other_n == 0){
    prob <- 100
    return(prob)
  }
  # 2.Y. don't go through loop if your lowest card is the card right after last card on pile
  else if (your_lowest == (last_pile+1)){
    prob <- 100
    return(prob)
    
  }
  # ----------------------------
  # 2.4. prob of lowest card, if there are no other possible lower cards for other player(s) to have
  #else if (length(possible_cards) == 0){
  #  prob <- 100
  #  return(prob)
  #}
  # ----------------------------
  
  # 2.5. prob of you having lowest card, if you know opponents only other card
  else if(length(possible_cards) == 1){
    prob <- ifelse(your_lowest < possible_cards, 100, 0)

    # ----------------------------
    #if (your_lowest < possible_cards){
    #  prob <- 100
    #} else if (your_lowest > possible_cards){
    #  prob <- 0
    #}
    # ----------------------------
    return(prob)
  }
  # 2.3. prob of lowest card, if other player(s) can have lower cards
  else {
    # counter for amount of time you had lowest card
    tally <- 0

    for (i in 1:sims_n){
      # simulate unknown cards
      # note: behavior of sample, when x is vector of length 1 (does c(1:x) not c(x))
      unknown_cards <- sample(x = possible_cards,
                              size = other_n,
                              replace = FALSE)
      
      # info to display, if testing
      if(print_iter == TRUE){
          print(paste0("i: ", i))
          print(c("your_hand (sorted):", sort(your_hand)))
          print(c("unknown_cards (sorted):", sort(unknown_cards)), collapse=" ")
          }
  
      # if your card lower than all other unknown cards, increment tally
      if (all(your_lowest < unknown_cards)){
        tally <- tally + 1
        
        # info to display, if testing
        if(print_iter == TRUE){
          print(all(your_lowest < unknown_cards))
          }
        }
  
      # info to display, if testing
      if(print_iter == TRUE){
        print(paste0("current_tally:", tally))
        print("-------------------------")
        }
      } # FOR LOOP ENDS
    }
  
  # 3. FINAL CALCULATIONS
  # probability your card being lowest: number of times it was lowest, out of all simulations
  prob <- tally/sims_n*100
  
  # info to display, if testing
  if(print_iter == TRUE){
    print(paste0("FINAL PROB: ", prob))
    print("~~~~~~~~~~~~~~~~~~~~~~~~~~")
  }
  
  # 4. FINAL OUTPUT
  # function's final output: probability your card was the lowest
  return(prob)
    
}

```


# 2. RUN FUNCTION

```{r}
mind_prob(your_hand = c(10, 66),
          last_pile = 7,
          other_n = 4,
          sims_n = 1000000,
          print_iter = FALSE)
```


# 3. GENERATE TABLE

The goal here is to create a table of probabilities of having the lowest card across these situations
- your_lowest: all cards from 1 to 100
- last card in pile: all cards from 0 to 100 (excluding your card)
- total cards in your hand: can be anything from 1 to 12
- total cards in other(s)' hands: can be anything from 0 to 36 (four max players, including your, round 12 = 12 cards each)

## 3.1. create data frame

```{r}
max_rows <- length(1:100) * length(0:100) * length(1:12) * length(0:36)

n_rows <- 0
for (your_lowest in 1:100){
  for (last_pile in 0:(your_lowest-1)){
    for (your_hand_size in 1:5){
      for (other_hand_size in 1:5){
        n_rows <- n_rows + 1
      }
    }
  }
}

prob_table <- 
  data.frame(your_lowest = numeric(n_rows),
             last_pile = numeric(n_rows),
             your_hand_size = numeric(n_rows),
             other_hand_size = numeric(n_rows),
             prob_you_lowest = numeric(n_rows))

max_rows
n_rows
prob_table
```


## 3.2. calculate table values

```{r}
row_i <- 0

for (your_lowest in 1:100){
  for (last_pile in 0:(your_lowest-1)){
    for (your_hand_size in 1:5){
      for (other_hand_size in 1:5){
        
        row_i <- row_i + 1

        if (your_hand_size == 1){
          rest_of_hand <- numeric(0)
          prob_i <-
             mind_prob(your_hand = your_lowest,
                      last_pile = last_pile,
                      other_n = other_hand_size,
                      sims_n = 1000,
                      print_iter = FALSE)
          }
        else if (your_hand_size > 1){
          probs_vec <- NA         
          for (hand_sim_i in 1:1000){
 
            rest_of_hand <- sample(x = setdiff(1:100, your_lowest),
                                   size = your_hand_size-1,
                                   replace = FALSE)
            probs_vec[hand_sim_i] <-
             mind_prob(your_hand = c(your_lowest, rest_of_hand),
                      last_pile = last_pile,
                      other_n = other_hand_size,
                      sims_n = 1000,
                      print_iter = FALSE)
          }
          prob_i <- mean(probs_vec, na.rm = TRUE)
         }
        
        prob_table[row_i, "your_lowest"] <- your_lowest
        prob_table[row_i, "last_pile"] <- last_pile
        prob_table[row_i, "your_hand_size"] <- your_hand_size
        prob_table[row_i, "other_hand_size"] <- other_hand_size
        prob_table[row_i, "prob_you_lowest"] <- prob_i

      }
    }
  }
}


# immediately save after running, since leaving on over night
# save as R data file
save(prob_table,
     file = "the_mind_prob_table_OVERNIGHT.Rda")

# also save as csv (for non R users)
write.csv(prob_table,
          file = "the_mind_prob_table_OVERNIGHT.csv")
```

## 3.3. view table

```{r}
prob_table
View(prob_table)
```

## 3.4.XXX. table misc.

```{r}
prob_table_1 <- prob_table[1:12000,]
View(prob_table_1)
```

## 3.5. save table

```{r}
# save as R data file
save(prob_table_1,
     file = "the_mind_prob_table.Rda")

# also save as csv (for non R users)
write.csv(prob_table_1,
          file = "the_mind_prob_table.csv")
```


# END/MISC

```{r}
mind_prob(your_hand = c(60, 70),
          last_pile = 44,
          other_n = 9,
          sims_n = 100000,
          print_iter = FALSE)
```

